### 2011-06-17 talk.tex
Talk final version

\documentclass{beamer}
\usepackage{stmaryrd}
\usepackage{amsmath}
\usepackage{graphicx}

\usetheme{Warsaw}

\newcommand{\etrans}[1]{\mathcal{E} \llbracket #1 \rrbracket}
\newcommand{\etranst}[1]{\mathcal{E}_t \llbracket #1 \rrbracket}
\newcommand{\etransf}[1]{\mathcal{E}_{f} \llbracket #1 \rrbracket}
\newcommand{\dtrans}[1]{\mathcal{D} \llbracket #1 \rrbracket}
\newcommand{\ttrans}[1]{\mathcal{T} \llbracket #1 \rrbracket}
\newcommand{\strans}[1]{\mathcal{S} \llbracket #1 \rrbracket}
\newcommand{\cf}[1]{\mbox{CF}(#1)}
\newcommand{\trans}[1]{\llbracket #1 \rrbracket}

\newcommand{\unr}{\texttt{UNR}}
\newcommand{\bad}{\texttt{BAD}}
\newcommand{\any}{\texttt{Any}}
\newcommand{\ok}{\texttt{Ok}}

\title{Contracts for Haskell}   
\author{Simon (PJ), Dimitrios, Koen, Charles-Pierre} 
\date{}

\begin{document}

\frame{\titlepage}

\section{Goals}  
\frame{
\frametitle{Goals}
We would like to write things like:\\
\ \\
\texttt{head ::: \{ x | not (null x) \} -> \ok\\
  head xs = case xs of\\
  | [] -> \bad \\
  | x:\_ -> x
}\\
\ \\ \ \\

\texttt{map ::: (Ok -> Ok) -> xs:Ok ->\\
  \quad \quad \quad \quad \quad \quad \{ ys | length ys = length xs \} }
\ \\ \ \\
Note: $\ok$ is just syntactic sugar for $\{ x \mid True \}$
}

\frame{
  \frametitle{Contract syntax}
  \begin{center}
  \begin{array}{rclr}
        c &:=& \{ x \mid e \}\\
        &\mid& \any \\
        &\mid& (c,c) \\
        &\mid& x{:}c_1 \to c_2\\
      \end{array}
  \end{center}
      \begin{itemize}
      \item Predicates should be any boolean Haskell expression.
      \item Everything (even errors!) satisfies $\any$\\
        \texttt{ fst ::: ($\ok,\any$) -> $\ok$ }
      \item Named arguments: $x$ is any value satisfying $c_1$.
      \end{itemize}
}

\frame{\frametitle{Crash-freeness}

  Morally, an expression $e$ is crash-free (cf) iff any \bad-free
  context in which we execute the expression in does not yield \bad.\\
  \ \\

  Laziness makes things tricky:
  \begin{itemize}
  \item \texttt{Just BAD}: not CF
  \item \texttt{(3,error ''foo'')}: not CF
  \item \texttt{fst (3,error ''foo'')}: \alert{CF}
  \end{itemize}
  \ \\
  And now to the operational semantics of contracts!
}

\frame{\frametitle{Contract semantics}
\begin{array}{rcl}
  e :::\{x \mid p \} &\iff& e \mbox{ diverges or } \\
  && (e \mbox{ is crash-free and } p[e/x] \not \to^\star \{\bad,False\})\\
  e ::: x:t_1 \to t_2 &\iff& \mbox{for all } e_1, e_1 ::: t_1 \mbox{ implies } (e~e_1) ::: t_2[e_1/x]\\
  e ::: (t_1,t_2) &\iff& e \mbox{ diverges or } \\
  && (e \to^\star (e_1,e_2) \mbox{ and } e_1 ::: t_1 \mbox{ and } e_2 ::: t_2)\\
  e ::: \any &\iff& True  
\end{array}
}

\frame{\frametitle{Contract checking}
  Two roads: runtime checking and static checking.
  \begin{itemize}
  \item Runtime: Findler \& Felleisen, many others ... 
  \item The road less taken: static checking
  \end{itemize}
  Our simple plan: 
  \begin{center}
    Haskell + Contracts $\to$ First-order logic $\to$ Automated Theorem Prover (equinox)
  \end{center}

  Equinox is Koen's automated theorem prover. It's a FO theorem
  prover with equality and is under active development.

}

\section{How we do it}

\frame{\frametitle{Our language}
  Ok, not really Haskell. Wlog we make these asumptions:
  \begin{itemize}
  \item Top-level only pattern-matching 
  \item $\lambda$-lifting (top-level function definitions)
  \item Saturated data constructors (l.o.g. here but only for the sake of clarity)
  \end{itemize}
  From now on, we only consider well-typed expressions.
}

\frame{\frametitle{Our syntax}
  \begin{array}{rcl}
    u,e &:=& x \mid f \mid e~e \mid D(e,\dots,e) \mid \bad\\
    mod &:=& def,mod \mid tdecl,mod \mid f ::: c,mod \mid \epsilon\\
    def &:=& f~x_1 \dots x_n = e \mid f~x_1 \dots x_n = \mbox{ case } e \mbox{ of } [(pat_i,e_i)]\\
    pat &:=& D(x_1,\dots,x_n)\\
    tdecl &:=& \mbox{data } T = dcons\\
    dcons &:=& \epsilon \mid D ; dcons \mid D ::: c ; dcons
  \end{array}
}

\frame{\frametitle{The big picture} Recall that $mod := def,mod \mid
  T,mod \mid f ::: c,mod \mid \epsilon$; a module is a list of
  function definitions, data definitions and claims that functions
  satisfy their contracts.
}

\section{The plan}
\frame{\frametitle{The big picture}
  \includegraphics[scale=0.4]{plan.pdf}
}

\frame{\frametitle{What we need}
  \begin{block}{Theorem (unproven)}
    $T_p \vdash T_c \implies f ::: c$
  \end{block}
  Unfortunately, so far we're only believers.

  Given this theorem we can use Koen's theorem prover to decide if
  $T_p \vdash T_c$.

}

\section{Translations}
\frame{\frametitle{Expressions become first-order logic terms}
  \begin{eqnarray*}
    \etrans{x} &=& x\\
    \etrans{f} &=& f\\
    \etrans{e_1~e_2} &=& app(\etrans{e_1},\etrans{e_2})\\
    \etrans{D(e_1,\dots,e_n)} &=& D(\etrans{e_1},\dots,\etrans{e_n})\\
    \etrans{\bad} &=& \bad
  \end{eqnarray*}
  \begin{itemize}
  \item Looks reasonable... But what's this ``app'' ?
  \item Why not $\etrans{e_1~e_2} = e_1(e_2)$ with $e_1$ being a FO function?
  \item To keep things first-order...
  \item Because we will need to quantify over an arbitrary
    expression (that includes functions)
  \end{itemize}
}

\frame{\frametitle{Formulae in the preamble}  
  We gave an (informal) definition of crash-freeness with the
  semantics of contract satisfaction. But here, we use CF as a logic
  predicate.

  We always add to our theory the following formulae, only
  approximating the intended semantics:
  $$ \forall f,x.~\cf{f} \land \cf{x} \implies \cf{app(f,x)} $$
  $$ \lnot \cf{\bad} $$
}

\frame{\frametitle{Translating modules}
  \begin{center}
    \begin{array}{rclr}
      \etrans{} &::& Expression \to Term\\
      \dtrans{} &::& Definition \to \{ FOF \}\\
      \ttrans{} &::& Data~type~declaration  \to \{ FOF \}\\
      \strans{} &::& Expression \to Contract \to \{ FOF \}\\
    \end{array}
  \end{center}

}

\frame{\frametitle{$\dtrans{}$ -- Definition translation (easy case)}
  Let $f$ be defined as $f~x = (x,x)$. It gives rise to the formula $\forall x.~app(f,x) = \mbox{Pair}(x,x)$.

  \begin{itemize}
  \item In the formula $f$ is an uninterpreted constant symbol.
  \item Pair is a function in FO logic.\\ \alert{We will give formulae for it soon!}
  \end{itemize}

  In general:
  $$\dtrans{f~x~y = e} = \forall x~y.~\etrans{f~x~y} = \etrans{e}$$
}

\subsection{Data constructors}
\frame{\frametitle{$\ttrans{}$ -- Data declaration translation}
  $\ttrans{\mobx{data } T = D_1 \mid \dots \mid D_n} = ?$

  Reasonable requirements:
  \begin{itemize}
  \item Injectivity
  \item CF conditions
  \end{itemize}

  Let's give some details...
}

\frame{\frametitle{$\ttrans{}$ -- Data constructor injectivity}
  $\ttrans{\mobx{data } T = D_1 \mid D_2} = ?$

  Assume the following arities: $D_1$:2, $D_2$:1.

  $$ \forall x_1,x_2 .~\forall y .~ D_1(x_1,x_2) \neq D_2(y) $$

  Do this for every pair of constructors!
}

\frame{\frametitle{$\ttrans{}$ -- Data constructor crash-freeness}
  $\ttrans{\mobx{data } T = D_1 \mid D_2} = ?$

  $$ \forall x_1,x_2.~\cf{x_1} \land \cf{x_2} \leftrightarrow \cf{D_1(x_1,x_2)} $$
  $$ \forall y.~\cf{y} \leftrightarrow \cf{D_2(y)}$$

  Values of type $T$ are crash-free iff they are created with crash-free values.

  \begin{itemize}
  \item $\rightarrow$ also true for cf function application
  \item $\leftarrow$  not true even for cf functions\\ (recall that \texttt{fst (3,$\bad$)} is crash-free)
  \end{itemize}
}

\subsection{Pattern matching}
\frame{\frametitle{$\dtrans{}$ -- The case of case}
  \texttt{
    head xs = case xs of\\
    ~~| Nil -> $\bad$\\
    ~~| Cons(y,ys) -> y
  }
  translates to:
  \begin{center}
  \begin{array}{rl}
    $       &app(head,Nil) = \bad\\
     \land &\forall y,ys.~app(head,Cons(y,ys)) = y\\
     \land & app(head,\bad) = \bad\\
     \land & \forall xs.~xs \neq Nil \land (\forall a~b.~xs \neq Cons(a,b))\\
     & ~~~\land xs \neq \bad \implies app(head,xs) = \unr$
  \end{array}
  \end{center}

  What is $\unr$? What is it for?
}
\subsection{Unreachability}
\frame{\frametitle{$\strans{}$ -- Contract satisfaction and the role of $\unr$}
  How do we translate that $e ::: c$?
  \begin{eqnarray*}
    \strans{e \in \any} &=& true\\
    \strans{e \in \{x \mid u \}} &=& e = \unr \lor (\mbox{CF}(\etrans{e}) \land \etrans{u[e/x]} \neq \bad\\
    &&~\land \etrans{u[e/x]} \neq False)\\
    \strans{e \in x:c_1 \to c_2}  &=& \forall x_1. \strans{x_1 \in c_1} \to \strans{e~x_1 \in c_2[x_1/x]}
  \end{eqnarray*}

  \begin{block}{$\unr$}
  Used to represent divergence or things that should not happen.

  $$fst ::: \ok \to \ok \mbox{ and }3 ::: \ok$$
  Thus, $fst~3$ should satisfy $\ok$!

  But it cannot happen (it's ill-typed), so we map $fst~3$ to a
  special value which satisfies $\ok$. (FOL is untyped!)
  \end{block}
}

\subsection{Recursion} 
\frame{\frametitle{$\dtrans{}$ -- Recursive functions: Take 1}
  Assume that we want to prove the claim $add ::: c$ where $c=\ok \to \ok \to \ok$.\\
  \texttt{add~x~y = case x of\\
    | Zero -> y\\
    | Succ a -> Succ (add a y)
  }\\

  Naively, we would do as usual: ask the theorem prover if
  $$\dtrans{add=e} \vdash \strans{add,c}$$
  But $add$ is recursive, so we would not go very far!

  \alert{Fix:} define a function $add\_rec$ that we use as IH. We now ask the theorem prover:
  $$\dtrans{add=e[add/add\_rec]},\strans{add\_rec,c} \vdash \strans{add,c}$$
}

\frame{\frametitle{$\dtrans{}$ -- Recursive functions: Take 1}
  That yields the following formula:
  \begin{center}
    \begin{array}{rl}
      $\forall x~y.&app(app(add,Zero),y) = y\\
      \land & \forall a.app(app(add,Succ(a)),y) = Succ(app(app(\alert{add\_rec},a),y))\\
      \land & app(app(add,\bad),y) = \bad\\
      \land & x \neq Zero \land (\forall c.~x \neq Succ(c))\\
      & ~~~\land x \neq \bad \implies app(app(add,x),y) = \unr$
    \end{array}
  \end{center}
  \ \\
  How about: $app(app(add,x),\bad) = \bad$?
}

\frame{\frametitle{$\dtrans{}$ -- Recursive functions: Take 2}
  A more complicated example: $mult ::: c$ where 
  $$c = \{x \mid x \neq 0\} \to \{y \mid y \neq 0\} \to \{z \mid z \neq 0\}$$

  \texttt{
    mult x y = case x of\\
    ~~| Zero -> Zero\\
    ~~| Succ a -> add (mult a y) y
    }

  $$\dtrans{mult=e[mult/mult\_rec]},\strans{mult\_rec,c} \not \vdash \strans{mult,c}$$
}

\frame{\frametitle{$\dtrans{}$ -- Recursive functions: Take 2}
  \texttt{
    mult x y = case x of\\
    ~~| Zero -> Zero\\
    ~~| Succ a -> add (mult a y) y
    }
  $$\dtrans{mult=e[mult/mult\_rec]},\strans{mult\_rec,c} \not \vdash \strans{mult,c}$$
  Why? Because the IH is too restrictive!\\
  \alert{Fix:}\\
  \begin{array}{rcl}
    $\dtrans{mult=e[mult/mult\_rec]}$&&\\
    $\strans{mult\_rec,c} & \vdash & \strans{mult,c}$\\\
    $\dtrans{mult\_rec=e[mult/mult\_rec]}$&&\\
  \end{array}
}

\frame{\frametitle{Questions}
  \begin{itemize}
  \item \emph{Proof that $T_p \vdash T_c \implies f ::: c$}
  \item Speed-up the theorem prover (by giving strategies)
  \item The user should be able to specify lemmas. How?
  \item Hoare Logic view of contracts (partial spec) vs. liquid types (complete spec)
  \item Contracts on datatypes
  \item Not all contracts expressible as Haskell predicates. Need special operator: \texttt{`satisfies` c}?
  \end{itemize}
}

\end{document}