### 2011-10-19 paper.tex
Paper: made a full comments pass through section 1.
Ouch!

\documentclass[preprint]{sigplanconf}

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{stmaryrd}
\usepackage[pdftex]{graphicx}

% Margin notes (ugly and complicated)
%% \usepackage[draft]{fixme}
%% \fxusetheme{color}
%% %\fxusetheme{signature}
%% %\fxuselayouts{pdfmargin}
%% %\fxusetargetlayout{color}
%% \FXRegisterAuthor{nc}{anc}{NC}

% Margin notes round two (easier and better looking)
%
% http://trackchanges.sourceforge.net/help_stylefile.html
\usepackage[inline]{trackchanges} % [margins] is also good
\addeditor{nc}

\begin{document}

\def\ruleform#1{{\setlength{\fboxrule}{1.2pt}\fbox{\normalsize $#1$}}}
\newcommand{\etrans}[1]{\mathcal{E} \llbracket #1 \rrbracket}
\newcommand{\dtrans}[1]{\mathcal{D} \llbracket #1 \rrbracket}
\newcommand{\ktrans}[1]{\mathcal{K} \llbracket #1 \rrbracket}
\newcommand{\ctrans}[1]{\mathcal{C} \llbracket #1 \rrbracket}
\newcommand{\trans}[1]{\llbracket #1 \rrbracket}

\newcommand{\tot}{\leftrightarrow}

\newtheorem{definition}{Definition}

\newcommand{\unr}{\texttt{UNR}}
\newcommand{\bad}{\texttt{BAD}}
\newcommand{\any}{\texttt{Any}}
\newcommand{\ok}{\texttt{Ok}}
\newcommand{\hprime}{\mathcal{H}'}
\newcommand{\cfc}{\texttt{CF}}
\newcommand{\cf}[1]{\texttt{CF}(#1)}
\newcommand{\weak}[1]{\mbox{$\$$weak}(#1)}
\renewcommand{\min}[1]{\mbox{min}(#1)}

\title{Static contract checking for Haskell}
\authorinfo{Simon, Koen, Dimitrios, Charles-Pierre}
\maketitle

\section{Introduction}

Our approach to static contract-checking is to translate source code
to a first-order logic theory and
 then use an automated theorem prover
to check the consistency of the theory.

Consider:\note[nc]{Unlike the POPL '09 paper, $CF(e)$ is not implied by $e \in \{x|p\}$.}
\begin{verbatim}
data List a = Nil | Cons a (List a)

notnull x = case x of
  Nil -> False
  Cons(x,y) -> True

head ::: (CF && {x | notnull x }) -> CF
head xs = case xs of
  Nil -> BAD
  Cons(x,y) -> x
\end{verbatim}

First, we need to encode the List structure.

We start by stating that $Nil$ and $Cons$ can never be equal:
\begin{equation*}
\forall a,b.~Cons(a,b) \neq Nil
\end{equation*}

Then, we must state that $Nil$ never crashes (ie cannot be evaluated
to an exception) and that $Cons(x,y)$ \add[nc]{never} crashes iff \change[nc]{either}{both} $x$ \change[nc]{or}{and} $y$
\add[nc]{never} crash. The statement $x$ \add[nc]{never} crashes is encoded by the term $\cf{x}$\note[nc]{Made more precise in section \ref{sec:cf}}.

\begin{equation*}
\cf{Nil}
\end{equation*}
\begin{equation}
  \label{cfcons}
\forall a,b.~\cf{a} \land \cf{b} \iff \cf{Cons(a,b)}
\end{equation}

We also say some stuff about \annote[nc]{unreachibility}{??? e.g. \[
  CF(fst(x,BAD)) \iff CF(x) \] as in section \ref{sec:cf}?} but I
can't think of a good way to explain it right now.
\begin{equation*}
\forall y,ys.~Cons(y,ys) \neq \unr 
\end{equation*}
\begin{equation*}
Nil \neq \unr 
\end{equation*}\note[nc]{\unr means ``infinite loop'', approximately (it's also the result of pattern matches on non-loop expressions with the wrong constructor head), and so these rules ($\Phi_4$) say that constructors are ``lazy'' or ``total''.}

Finally, we define projections for $Cons$. \annote[nc]{It is not
  strictly necessary, but it will be handy:}{??? Why is it handy? It
  doesn't seem to play any role in formal part of this paper (besides
  being specified in axioms $\Phi_1$ (projections)). In the current
  implementation it's used to avoid introducing existentially
  quantified variables in the $\bigvee_i$ part of $\dtrans{}$ for
  functions defined by case expressions. Equation \eqref{unrnotnull}
  below uses it that way. ??? Does Equinox care?}
\begin{equation*}
\forall xs,y,ys.~sel_{1,Cons}(Cons(y,ys)) = xs \implies xs = y 
\end{equation*}
\begin{equation*}
\forall xs,y,ys.~sel_{2,Cons}(Cons(y,ys)) = xs \implies xs = ys 
\end{equation*}\note[nc]{??? Why not the simpler \[\forall y,ys.~sel_{2,Cons}(Cons(y,ys)) = ys\]? Same question in Figure \ref{dtrans}.  (Equinox does have a built in notion of equality, e.g. running equinox on
\center{\texttt{fof(\_,theorem,? [X,Y] : (X = Y \& f(X) != f(Y))).}}\\
 results in ``Unsatisfiable'', so support for equality reasoning is not the reason.)}
%% \[
%% \text{\texttt{\$ echo "fof(\_,theorem,? [X,Y] : (X = Y & f(X) != f(Y)))."}} \\
%%                  \text{\texttt{| equinox /dev/stdin}}}}
%% \]

Now we translate the $null$ function. Note that the symbols $true$ and
$false$ are the representation of the data constructors $True$ and
$False$ in Haskell, not the boolean values $\top$ and $\bot$ in our
logic.\note[nc]{XXX, TODO: use $True$ and $False$ in the
  text.  The lowercase $true$ and $false$ is an artifact of the implementation, i.e. Equinox accepts TPTP format, which doesn't allow bare uppercase function symbols.  Moreover, the lower case here is inconsistent with the rest of the presentation, e.g. $Cons$ and $Nil$ above.}
\begin{equation*}
\forall xs.~ xs=Nil \implies notnull(xs) = false
\end{equation*}\note[nc]{Here we might ask the same question as above: why not use the simpler definition
\[
\forall notnull(Nil) = false
\]? The answer is that this is an expansion of the $etrans{e} = K_i(\vec{x_i}) \to a = \etrans{e_i}$ part of the $\dtrans{}$ rule for functions defined by cases.  I.e., the scrutinee $Nil$ is, in general, an expression involving, but not necessarily equal to, the arguments to the function.  In this special case we could remove the implication, but not in general.  The $a = \etrans{e_i}$ part is missing above, but I guess that's because this text was not updated for the $\min$-translation.}

\begin{equation}
  \label{defnotnull}
\forall xs,y,ys.~ x=Cons(y,ys) \implies notnull(xs) = true
\end{equation}

We also need to specify the translation of calls to $notnull$ with
$\bad$ values (to encode the fact that $notnull(\bad) = \bad$.
\begin{equation*}
\forall xs.~xs=\bad \to notnull(xs) = \bad
\end{equation*}

Finally, we say that one call $notnull$ with a an argument which is
not $Nil$ or $Cons$ or $\bad$ then the result is $\unr$.
\begin{align}
  \label{unrnotnull}
  \forall xs.~xs \neq \bad \land xs \neq Nil  \notag \\
  \land xs \neq Cons(sel_{1,Cons}(xs),sel_{2,Cons}(xs)) \to  \notag \\
  notnull(xs) = \unr 
\end{align}

The translation of $head$ follows the same pattern:
\begin{align}
  \forall xs.~ x=Nil \implies head(xs) = \bad \notag\\
  \forall xs,y,ys.~ xs=Cons(y,ys) \implies head(xs) = y     \label{defhead} \\
    \forall xs.~xs=\bad \to head(xs) = \bad \notag
\end{align}\note[nc]{??? Why not use the same project+construct trick from the last equation?  I.e., why not
\[
\forall xs.~ xs=Cons(sel_{Cons,1}(xs),sel_{Cons,2}(xs)) \implies head(xs) = sel_{Cons,1}(xs)
\]?  Because it's ugly?  Would Equinox care?}

\begin{align*}
 \forall xs.~xs \neq \bad \land xs \neq Nil \notag \\
 \land xs \neq Cons(sel_{1,Cons}(xs),sel_{2,Cons}(xs)) \to  \notag \\
 head(xs) = \unr 
\end{align*}

We now have translated the source code. Let us call all those formulae
the theory $T$. We translate separatly the contract:
\begin{equation*}
  \phi := \forall xs.~\cf{xs} \land notnull(xs) = true \to \cf{head(xs)} 
\end{equation*}

Now that we translated everything to first-order logic, we can ask the
theorem prover if the theory formed by those formulae \change[nc]{is
consistent}{entails the contract}, ie if $T \vdash \phi$.

Intuitively, $T$ is consistent (ie $T \not \vdash \bot$), because each
formula serves a specific purpose.\note[nc]{

  A more precise consistency argument by exhibiting a model for $T$:
  in the model we define $\min{}$ to be the trivial predicate,
  i.e. $\forall x. \min{x}$.  So, we can drop all the $\min{}$'s.  The
  model for the resulting formulas is a language with the syntax used
  in this paper ($\mathcal{H}'$ below).

  The carrier consists of all closed (i.e. variable free / grounded)
  programs that can be written using the functions and constructors in
  $T$, using the expression syntax of $\mathcal{H}'$ below, and
  evaluated as described in the following paragraphs.  We assume $T$
  comes from a ``well-typed'' source program $P$, i.e. s.t. the
  Haskell program corresponding to $P$ is well-typed.

  The semantics are what you would expect for ``well-typed''
  terminating expressions (again, ``well-typed'' in the sense that the
  corresponding Haskell expression is well-typed in an environment
  including $P$'s decls). For example, \texttt{Cons(Z,Nil)} is
  ``well-typed'' but \texttt{Cons(Nil,Z)} is ``ill-typed'' in a the
  context of a program declaring List and Nat types.

  In general, scrutinizing a \unr{} yields \unr, as does scrutinizing
  a terminating expression with the wrong constructor head
  (e.g. \texttt{tail(S Z) =} \unr{}). Scrutinizing an ill-typed
  expression with a constructor head of the right type evaluates as
  specified by the pattern (e.g. \texttt{tail(Cons(Nil,Z)) = Z}),
  unless there is no corresponding pattern (``non-exhaustive
  pattern''), in which case the result is \bad{}.  NB: wrong-type and
  no-corresponding-pattern are treated very differently.  The
  intuition is that the former never happens in a well-typed program.

  Applying \unr{} or \bad{} as a function results in \unr{} or \bad{},
  respectively.

  Term constructors have disjoint ranges, and are total.

  I think that covers functions and datatypes, but it remains to argue
  that the axioms for contracts are consistent \ldots.

  \ldots but maybe this is tricky?  Modeling $CF(e)$ seems easy,
  i.e. just ask the oracle used above for evaluation if $e$ is \bad{}
  or not.  But recursion seems problematic, since we get to assume in
  $T$ that the contract holds in the recursive position (Section
  \ref{sec:recursiveContracts}), and so it seems we need to argue that
  \emph{all} expressable contracts are consistent.

  Here's a flawed contradiction ``proof'': consider
  \[
  e = e ::: CF \&\& \neg CF.
  \]
  Now, $T$ will contain $CF(ep) \land \neg CF(ep)$, and were hosed.
  This is flawed because there is no $\neg$ in the annotated
  contracts.

  OK, and actually this might be easy: we simply interpret all
  recursive function symbols (the $fp$'s for recursive $f$'s) as
  \unr{}, and use the fact that forall $x$, $\unr{}~x = \unr$.  Now,
  an easy induction on contract translations (Figure \ref{ctrans})
  gives \[ \emptyset \vdash \ctrans{\unr \in c}
  \]
  for all contracts $c$, and we're done!  Well, this is really just
  the base case, where all contracts in $T$ are for otherwise
  undefined recursion symbols ($fp$'s), but the inductive case is just
  as easy: given a consistent $T$, which may include contracts for
  defined symbols, we can always extend it to a consistent $T'$ by
  adjoining constracts for undefined symbols.

}

Now, assume that $xs$ satisfies
$\cf{xs}$ and $notnull(xs) = true$.  We can derive that $xs \neq \bad$
because we have $\cf{xs}$ and \annote[nc]{$\lnot \cf{\bad}$}{Obviously true, given the definition of $CF$, but it's worth noting that this plays a key part in the axiomitization of $CF$, which is ommited in this draft (so far ...)}. The constraint
$notnull(xs) = true$ doesn't directly imply that $xs = Cons(y,ys)$ for
some $y$ and $ys$. But $notnull$ is totally defined, because of
\eqref{unrnotnull}. This implies (by \eqref{defnotnull}) that there exist
$y$ and $ys$ such that $xs = Cons(y,ys)$. Recalling $\cf{xs}$, we can
now derive $\cf{y}$ and $\cf{ys}$ (by \eqref{cfcons}). But $head(xs) =
y$ because of \eqref{defhead}, and $y$ is crash-free, so we can finally
derive $\cf{head(xs)}$. QED.

\section{Languages}
\subsection{$\hprime:~\lambda$-calculus variant}

The syntax of $\mathcal{H}$ is defined in figure~\ref{hprime-stx}. A module
is a list of toplevel definitions, claims that functions statisfy
contracts and data definitions.

\begin{itemize}
\item There's no $\lambda$-abstraction, because we can always lift
  them to toplevel declaration.
\item We do not allow nested case expressions, because once again, we
  can always lift them to the toplevel.
\item Until section~\ref{ho} we will only consider full application of
  functions ($f(x,y)$), in order to remove clutter. Dealing with
  partial application is not hard but a bit cumbersome.
\end{itemize}

\begin{figure}[h]
  \centering
  \[
  \begin{array}{rclr}
    mod &:=& def_1,\dots,def_n &\\
    def &\in& \mbox{Definition} & \\
    def &:=& \texttt{data } T = K_1 a_1 \mid \cdots \mid K_n a_n& \mbox{Data definition}\\
    &\mid& f \in c & \mbox{Contract claim}\\
    &\mid& f~\vec{x}~=~e & \\
    &\raisebox{ 0.18 cm }{$\mid$}& \multicolumn{2}{c}{
      \begin{array}{rcl}
        \hspace{-0.2cm} f~\vec{x} &=& \texttt{case } e \texttt{ of } \\
         && K_1(\vec{x}_1) \to e_1 \mid \cdots \mid K_n(\vec{x}_n) \to e_n \\
       \end{array} 
       } \\
  \end{array} \]

  \[  \begin{array}{rclr}
    x,y,f,g,a,b & \in & \mbox{Variables} \\
    T &\in& \mbox{Type Constructors} \\
    K &\in& \mbox{Data Constructors} \\
  \end{array} \]

  \[  \begin{array}{rclr}
    e &\in& \mbox{Expressions} & \\
    e &::=& x & \\
    &\mid& \bad & \\
    &\mid& e~e & \\
    &\mid& f(e,\dots,e) & \\
    &\mid& K(e,\dots,e) & \\
  \end{array} \]

\caption{Syntax of the language $\hprime$}
\label{hprime-stx}
\end{figure}

\subsection{Contracts}
Contract syntax is described in figure \ref{cont-stx}. The predicates
we use in our contracts can be any boolean $\hprime$ expression. We
only consider pairs of contract for simplicity, although there is no
issue with generalisation to arbitrary tuples.

\begin{figure}[h]
 \centering 
  \[  \begin{array}{rclr}
  c &:=& x:c \to c\\
  &\mid& (c,c) \\
  &\mid& c \land c \\
  &\mid& c \lor c \\
  &\mid& \{ x \mid p \}\\
  &\mid& \cfc \\
  \end{array} \]
  \caption{Contract syntax}
  \label{cont-stx}
\end{figure}

We give the semantics of contract by defining ``$e$ satisfies $t$'',
written $e \in t$ in figure \ref{cont-smt}. Note that this definition
doesn't yield any operative way to check that an expression actually
meets the specification given by its contract.

\begin{figure}[h]
 \centering
  \[  \begin{array}{rcl}
    e \in \{x \mid p \} &\iff& e \mbox{ diverges or } p[e/x] \not \to^\star \{\bad,False\}\\
    e \in x:c_1 \to c_2 &\iff& \forall e_1 \in c_1, (e~e_1) \in c_2[e_1/x]\\
    e \in (c_1,c_2) &\iff& e \mbox{ diverges or }\\
    &&  (e \to^\star (e_1,e_2) \mbox{ and } e_1 \in c_1, e_2 \in c_2)\\
    e \in c_1 \land c_2 &\iff& e \in c_1 \mbox{ and } e \in c_2 \\
    e \in c_1 \lor c_2 &\iff& e \in c_1 \mbox{ or } e \in c_2 \\
    e \in \cfc &\iff& e \mbox{ is crash-free}
  \end{array} \]
  \caption{Semantics of contract satisfaction}
  \label{cont-smt}
\end{figure}

\subsection{Crash-freeness}\label{sec:cf}
Note that $\cfc$ represents two things: it can be a contract, as in $f
\in \cfc$ or a special formula in first-order logic $\cf{f}$.

We use $\bad$ to signal that something has gone wrong in the program :
it has crashed.

\begin{definition}[Crash]
A closed term $e$ crashed iff $e \to^* \bad$.
\end{definition}

\begin{definition}[Diverges]
A closed expression $e$ diverges iff either $e \to^* \unr$ or there is
no value $val$ such that $e \to^* val$
\end{definition}

\begin{definition}[Syntactic safety]
A (possibly open) expression $e$ is syntactically safe iff $\bad \not
\in_s e$. Similarly a context \cal C is syntactically safe iff $\bad
\not \in_s \mathcal{C}$.
\end{definition}

The notation $\bad \not \in e$ means that $\bad$ does not appear
anywhere in $e$, similarly for $\bad \not \in_s \mathcal{C}$. For
example, $Just 3$ is syntactically safe whereas $Just \bad$ is not.

\begin{definition}[Crash-free]
An expression $e$ is said to be crash-free iff 
$$\forall \mathcal{C}.~\bad \not \in_s \mathcal{C} \mbox{ and } \vdash
\mathcal{C} \llbracket e \rrbracket :: () \mathcal{C} \not \to^* \bad$$
\end{definition}
The notation $\mathcal{C} \llbracket e \rrbracket :: ()$ means that
$\mathcal{C} \llbracket e \rrbracket$ is closed and well-typed.  Note
that there are crash-free expression that are not syntactically safe,
for example $fst~(1,\bad)$.

\subsection{BAD and UNR}
Consider the following piece of code:
\begin{verbatim}
a = 0 + True

b ::: CF
b = undefined

c = error "foo"
\end{verbatim}

\begin{itemize}
\item $a$ is ill-typed
\item $b$'s implementation is not correct wrt its contract
\item $c$ goes through the whole toolchain (compiler, typechecker, contractchecker)
\end{itemize}

One thing to notice is that $a$ and $b$ are things that ``sould not
happen'' but are caught statically whereas $c$ should not happen but
can only be dealt with dynamically.

We can now define two types of problematic expressions: those that
cannot happen during a run of the program and those that can.
Expressions of the first type are called unreachable (and equated to
the special value $\unr$ in our first-order theory), whilst
expressions of the latter type are called bad (and equated to the
special value $\bad$).

We said earlier that we only considered syntactically correct and
well-typed programs as input. That implies that the ``$a$'' case
cannot happen. But given that our first-order logic is not typed, the
theorem prover may decide to instanciate a variable with an ill-typed
value! In order to prevent this, we will need to encode some basic
type-checking mecanism directly in our first-order theory.

\subsection{First-order logic with equality}
We use first-order logic with equality, defined in figure \ref{fol-stx}.

\begin{figure}
 \centering
  \[  \begin{array}{rcl}
    v,w,s,t &:=& x \mid K(t,\dots,t) \mid f(t,\dots,t) \mid app(t,t)\\
    && \mid \bad \mid \unr \\
    \phi &:=& \forall x.\phi \mid \lnot \phi \mid \phi \lor \phi \mid \top \mid \bot \mid t=t \mid \mbox{CF}(t)\\
    && \mid \phi \land \phi \mid \phi \to \phi \mid \phi \tot \phi \\
    \Phi &:=& \epsilon \mid \phi \mid \Phi \cup \Phi \\
  \end{array} \]
  \caption{First-order logic syntax}
  \label{fol-stx}
\end{figure}

\section{Translations}
For an overview of the different translations we define, see figure~\ref{trans}

\begin{figure}
 \begin{center}
  \[  \begin{array}{rcl}
    \etrans{\mbox{expression}} &\to& t~(\mbox{Term}) \\
    \dtrans{def} &\to& \Phi~(\mbox{Set of formulae}) \\
    \ktrans{\mbox{data } T = \dots} &\to& \Phi~(\mbox{Set of formulae}) \\
    \ctrans{f \in c} &\to& \Phi~(\mbox{Set of formulae}) \\
  \end{array} \]
  \end{center}
  \caption{Translations}
  \label{trans}
\end{figure}

\subsection{$\etrans{}$ -- Expressions}
Our most basic translation is from expressions in $\hprime$ to terms
in first-order logic. Given this translation we will be able to
translate definitions, data types and contracts to first-order
formulae. It is described in figure~\ref{etrans}.

\begin{figure}
$$
\begin{array}{rcl}
\multicolumn{3}{c}{\ruleform{\etrans{e} = t}} \\[2.5mm]
 \etrans{x} &=& x\\
  \etrans{f(e_1,\dots,e_n)} &=& f(\etrans{e_1},\dots,\etrans{e_n})\\
  \etrans{K(e_1,\dots,e_n)} &=& K(\etrans{e_1},\dots,\etrans{e_n})\\
  \etrans{\bad} &=& \bad
\end{array}
$$
\caption{$\etrans{}$ -- Expression translation}
\label{etrans}
\end{figure}

\subsection{$\dtrans{}$ -- Definitions}
We give in figure~\ref{dtrans} the two translations of function
definitions.

figure~\ref{dtrans-reg} gives the translation of function not defined by
pattern matching, which is really easy: we just have to state the
equality between le left-hand side and the right-hand side.

Translating definitions that use pattern-matching is more challenging
and is described in figure~\ref{dtrans}.

The first line says that when applied to an argument that matches a
pattern of the case expression, we should equate the function call to
the corresponding expression.

The second line states that if the pattern-matching failed or if we
pattern-matched on $\bad$ then the result should be $\unr$.

\begin{figure*}
\[
\begin{array}{rcl}
  \multicolumn{3}{c}{\ruleform{\dtrans{def} = \Phi}} \\[2.5mm]
  \dtrans{\mbox{data } T = K_1, \dots, K_n} &=& \ktrans{\mbox{data } T = K_1, \dots, K_n} \\
  \dtrans{f \in c} &=& \ctrans{f \in c}\\
  \dtrans{f(\vec{x}) = e} &=& \forall \vec{x},a.~ (\min{a} \land a =
  f(\vec{x})) \to a = \etrans{e} \\
   \raisebox{ 0.5 cm }{$\dtrans{f(\vec{x}) = \mbox{case } e \mbox{ of } [K_i(\vec{x_i}) = e_i]}$} &\raisebox{0.5cm}{=}&  \hspace{-0.18cm} \begin{array}{rclcl}
     \forall \vec{x},a.~(\mbox{min}(a) \land a = f(\vec{x})) &\to&
     \multicolumn{3}{c}{  \texttt{min}(e) \land (e = \texttt{BAD} \lor
       \left(\bigvee_i \exists \vec{x_i}.~e = K_i(\vec{x_i})\right) \lor a = \texttt{UNR})} \\
     &\land& \forall \vec{x_1}.~\etrans{e} = K_1(\vec{x_1}) &\to& a = \etrans{e_1} \\
     &\land& & \cdots & \\
     &\land& \forall \vec{x_n}.~\etrans{e} = K_n(\vec{x_n}) &\to& a = \etrans{e_n}} \\
   \end{array} 
\end{array}
\]
\caption{$\dtrans{}$ -- Defintions translation}
 \label{dtrans}
 \note[nc]{??? Why do we introduce the extra quantified variable $a$,
   representing the result (``answer''?), instead of just using the result directly
   (which is logically equivalent)? Same question for the next
   Figure.  Does this make a difference to Equinox?}
\end{figure*}

\subsection{$\ktrans{}$ -- Datatypes}
We break down the translation for datatypes in four parts, described in figure~\ref{ktrans}
$$\ktrans{\mbox{data } T = K_1, \dots, K_n} = \Phi_1 \cup \Phi_2 \cup \Phi_3 \cup \Phi_4$$

\begin{itemize}
\item $(\Phi_1)$ For each $K_i$ of arity $a_i$ we introduce selectors
  $sel_{k,K_i}$, which are the projection of $K_i(x_1,\dots,x_{n_i})$
  on its $k$-th component.
\item $(\Phi_2)$ For each pair of constructors $K_i,K_j$, we state that they can
  never map to the same value.
\item $(\Phi_3)$ Then, we have to give crash-freeness conditions for each $K_i$:
  Notice that we have a equivalence.
  \begin{itemize}
  \item $\leftarrow$: if we pack crash-free values in a data
    constructor, the resulting value is crash-free.
  \item $\rightarrow$: a value $t$ of type $T$ is crash-free implies
    that every value packed in it is crash-free. Recall that one can
    define projection on any argument of a value of type $t$. So if
    the k-th argument of $t$ is not crash-free, then the k-th
    projection is a crash-free context that throws an expression that
    is not crash-free.

  Note that this is not true for functions: a function is not required
  to use all of its arguments. \texttt{fst} is crash-free if and only
  if the first argument of the pair is crash-free. The second argument
  being crash-free or not doesn't matter.
  \end{itemize}

\item $(\Phi_4)$ None of the $K_i$ is unreachable.
\item One may want to also state that if $\vec{x} \neq \bad$ then
  $K_i(\vec{x}) \neq \bad$. It is already implied by the fact that
  $\cf{\vec{x}} \to \cf{K_i(\vec{x})}$.
\end{itemize}

\begin{figure*}
  $$ \ruleform{\ktrans{data~def} = \Phi} $$
$$\ktrans{\mbox{data } T = K_1, \dots, K_n} = \Phi_1 \cup \Phi_2 \cup \Phi_3 \cup \Phi_4$$
 \hspace{5.6cm}where
  \begin{center}
    \[  \begin{array}{rcl}
      \Phi_1 &=& \bigcup_{1 \leq i \leq n}~\forall \vec{x},a.~(\min{a} \land a = K_i(\vec{x})) \to \bigwedge_{1 \leq j \leq k} x_j = sel_{j,K_i}(a) \\
      \Phi_2 &=& \bigcup_{1 \leq i < j \leq n}~\forall \vec{x},\vec{y},a.~ \lnot (\min{a} \land a = K_i(\vec{x}) \land a = K_j(\vec{x})) \\
      \Phi_3 &=& \bigcup_{1 \leq i \leq n}~\forall \vec{x},a.~(\min{a} \land a = K_i(\vec{x})) \to ((\bigwedge_{1 \leq j \leq k} \cf{x_j}) \tot \cf{a\text{\note[nc]{This was $K_i(\vec{x})$.  Typo, right?}}}) \\
      \Phi_4 &=& \bigcup_{1 \leq i \leq n}~\forall \vec{x},a.~(\min{a} \land a = K_i(\vec{x})) \to a \neq \unr \land a \neq \bad 
    \end{array} \]
  \end{center}
  \caption{$\ktrans{}$ -- Data type translation}
  \label{ktrans}
\end{figure*}

\subsection{$\ctrans{}$ -- Contracts}
We give in figure~\ref{ctrans} the translation of contract satisfaction.
$true$ refers to the translation to a term of the data constructor
\texttt{True} in $\hprime$, not to the actual true value.

Note that we define the translation of $f \in c$ and of $f \not \in
c$. We have to do that because one is not the negation of the other,
even though $\lnot (f \not \in c)$ implies $f \in c$.

\begin{figure*}
  $$ \ruleform{\ctrans{contract} = \Phi} $$
\begin{eqnarray*}
  \ctrans{e \in \{x \mid b \}} &=&  \min{\etrans{b[e/x]}} \land (\etrans{b[e/x]} = true \lor \etrans{e} = \unr)\\
  \ctrans{e \in x:c_1 \to c_2} &=& \forall x.~\min{\etrans{e~x}} \to (\ctrans{x \not \in c_1} \lor  \ctrans{e~x \in c_2})\\
  \ctrans{e \in \cfc} &=& \min{\etrans{e}} \land \cf{\etrans{e}}
\end{eqnarray*}

\begin{eqnarray*}
  \ctrans{e \not \in \{x \mid b \}} &=&  \min{\etrans{b[e/x]}} \land (\etrans{b[e/x]} = false \lor \etrans{e} = \bad)\\
  \ctrans{e \not \in x:c_1 \to c_2} &=& \exists x.~ \ctrans{x \in c_1} \land  \ctrans{e~x \not \in c_2}\\
  \ctrans{e \not \in \cfc} &=& \min{\etrans{e}} \land \lnot \cf{\etrans{e}}
\end{eqnarray*}

\begin{center}
\[  \begin{array}{rcl}
 \ctrans{e \in c_1 \&\& c_2} &=& \ctrans{e \in c_1} \land \ctrans{e \in c_2} \\
 \ctrans{e \not \in c_1 \&\& c_2} &=& \ctrans{e \not \in c_1} \lor \ctrans{e \not \in c_2}\\
 \ctrans{e \in c_1 || c_2} &=& \ctrans{e \in c_1} \lor \ctrans{e \in c_2} \\
  \ctrans{e \not \in c_1 || c_2} &=& \ctrans{e \not \in c_1} \land \ctrans{e \not \in c_2}\\
 \ctrans{(a,b) \in (c_1,c_2)}   &=& \ctrans{a \in c_1} \land \ctrans{b \in c_2} \\
 \ctrans{(a,b) \not \in (c_1,c_2)}   &=& \ctrans{a \not \in c_1} \lor \ctrans{b \not \in c_2} \\
\end{array} \]
\end{center}

\caption{$\ctrans{}$ -- Contract translation}
\label{ctrans}
\end{figure*}

\section{$\trans{}$ -- Checking a module}

\subsection{Prelude}
There are some formulae that should always be included in our FO theory.

We need to state that $\bad$ is not crash-free with the formula:
$\lnot \cf{\bad}$.

Plus we need to give formulae for the boolean datatype and for
 unreachability. Strictly speaking, we can omit them and just add the
following lines to source files:
\begin{verbatim}
data UNR = UNR
data Bool = True | False
\end{verbatim}
But given that those datatypes are used by our translation, we can
just directly include their translation every time we translate a
module.

\subsection{Contract checking -- Non-recursive case}
Input: a module $M$ that consists of a list of definitions, datatypes,
contracts and a contract $c$ for a non-recursive function $f$ this is
defined in $M$.

We say that the function implementation is correct wrt to its contract
iff $$\trans{M} \vdash \ctrans{f \in c}$$

\subsection{Contract checking -- Recursive case}\label{sec:recursiveContracts}
If the function $f=e$ is recursive, then we ask the theorem prover the
following:

$$\trans{M - f}, \dtrans{f=e[f/f_p]}, \ctrans{f_p \in c} \vdash \ctrans{f \in c}$$

Where $M - f $ means the content of the module $M$ without $f$'s
definition and $f$'s contract. TODO Stress that it's not always enough
and that we may have to unroll several times!

\subsection{Module checking}
A module is a collection of function definitions, data definitions and
contracts. What we want to do is to check that functions satisfy their
contract(s).

\subsubsection{Naive example}
Here is a little example showing that we should be careful about which
formulae should belong to a theory.

Assume that we have a module that contains two functions defintion $f$
and $g$ and two contracts : $f \in \cfc$ and $g \in \cfc$. We assume
that those contracts do not hold, for example if $f$ is $head$ and $g$
is $last$.

First, we want to check $f$'s contract. So we ask the theorem prover
if
$$ \dtrans{f}, \dtrans{g}, \ctrans{g} \vdash \ctrans{f} $$

But, given that $g$'s contract does not hold, we can derive $\bot$ and
then prove that $f$'s contract hold.

For the same reason, we can prove that $g$ contract's holds, when in
fact it doesn't.

Finally, the user thinks he's done, but in fact he has proven nothing.

\subsubsection{The proper way to check a module}
Consider the following situation, where $a$'s definition relies on $f$
and $g$.
\begin{center}
\includegraphics[scale=0.3]{flow.png}
\end{center}

We should only include formulae that belongs to functions that are
actually used. For example, to prove $a$'s contract, we should only
include $f$ and $g$ translations, and we would ask equinox:
$$ \dtrans{f},\dtrans{g},\dtrans{a},\ctrans{f},\ctrans{g} \vdash
\ctrans{a}$$

\section{Correctness of the translation}
For the translation to be useful, $\trans{M} \vdash \ctrans{f \in c} $
should imply that $f \in c$.

\section{Higher-orderness}
\label{ho}
Our current translation only considers fully applied functions and
first-order functions. For example, so far, we cannot give any
contract to $map$ because it would involve quantifying over a
function, a thing that is not first-order.

There is a possible workaround, which involves the ``app'' function
defined in our first-order logic. Assume we have a function $f$ that
is not fully applied somewhere in a module. We create the term
$f\_ptr$ which relates to $f$ by the equations given in figure~\ref{ho-fig}

This way, we can emulate quantification over function by quantifying
on their $ptr$ counterpart.

\begin{figure*}
$$\etrans{e_1~e_2} = app(e_1,e_2)$$
$$ \forall x_1,\dots,x_n.~f(x_1,\dots,x_n) = app(app(\dots app(f\_ptr,x_1),x_2),\dots,x_n)$$
$$ \cf{f\_ptr} \tot \forall x_1,\dots,x_n.~\cf{x_1} \land \dots \land \cf{x_n} \to \cf{f(x_1,\dots,x_n)}$$
$$\forall f\_ptr,x.~\cf{f\_ptr} \land \cf{x} \to \cf{app(f\_ptr,x)}$$
\caption{Encoding of higher-orderness}
\label{ho-fig}
\end{figure*}

\section{Experiments}
That's how we roll.

\begin{figure*}
\begin{center}
    \begin{tabular}{l|c|c|c|c|c}
      Problem & Equinox & Equinox (+ weak) & SPASS & Vampire & E \\
      \hline
      Add.hs & 0.25 & 0.08 & 0.04 & 0.12 & 0.05\\
      BinaryTree.hs & 0.45 & 0.2 & 0.04 & 0.01 & 0.04 \\
      Branch.hs & 0.27 & 0.40 & 0.04 & 0.01 & 0.03 \\
      Copy.hs & 0.86 & 0.09 & 0.03 & 0.01 & 184.3 \\
      Head.hs & 0.32 & 0.29 & 0.03 & 0.03 & 4.2 \\
      Implies.hs & 3.24 & 0.32 & 0.06 & 0.02 & 0.11 \\
      Map.hs & 2.47 & 0.14 & 0.92 & 1.02 & $>$300 \\
      Mult.hs & $>$300 & 0.41 & 0.05 & 0.22 & 11.71 \\
      Multgt.hs & $>$300 & 1.24 & 0.62 & 1.31 & $>$300 \\
      NatEq.hs & 203.12 & 0.33 & 0.02 & 0.03 & 0.343 \\
      Odd.hs & 0.42 & 1.17 & 0.06 & 0.03 & $>$300 \\
      Reverse.hs & 72.32 & 0.12 & 0.05 & 0.02 & 0.038 \\
      Simple.hs & 0.07 & 0.04 & 0.01 & 0.01 & 0.022 \\
      Test.hs & 7.76 & 2.86 & 0.08 & 0.05 & $>$300 \\
      Test2.hs & 5.63 & 0.09 & 0.07 & 0.01 & 1.02 \\
    \end{tabular}
\end{center}
\caption{Comparison (in seconds) with other theorem provers}
\label{comparison}
\end{figure*}

\end{document}